A longitudinal analysis of the social information in infants' naturalistic

### 2022-01-07 paper/manuscript/appendix.Rmd
save appendix pdf figs in line

```{r handLocation, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=6.5, fig.height=3.5, fig.cap = "Density estimates for the child (left) and adult (right) hands that were detected in the 24K frame random gold set; each dot represents the center of a bounding box made by an adult participant. Brighter values indicate more detections."}
cowplot::plot_grid(child_hand_plot, adult_hand_plot, nrow=1)
```

@@

# density plot of OP's false positive face/hand locations
p1 <- ddd %>% filter(face_eval_ketan=="FP", label=="face") %>%
  mutate(x=left+0.5*width,
         y=1-(top+0.5*height)) %>% # could actually overplot all the face rectangles? prob very messy...
  ggplot(aes(x=x, y=y, fill=..level..)) + 
  stat_density_2d(geom = "polygon") + xlim(-.05,1.05) + ylim(-.05,1.0) + 
  xlab("Center of OpenPose Detection") + 
  ylab("Center of OpenPose Detection") + 
  ggtitle("Face False Positives") + theme(legend.position = "none") +
  scale_y_reverse() +
  theme_few(base_size=12)

p2 <- ddd %>% filter(hand_eval_ketan=="FP", label=="hand") %>%
  mutate(x=left+0.5*width,
         y=1-(top+0.5*height)) %>% # top-left is 0,0
  ggplot(aes(x=x, y=y, fill=..level..)) + 
  stat_density_2d(geom = "polygon") + xlim(-.05,1.05) + ylim(-.05,1.0) + 
  xlab("Center of OpenPose Detection") + 
  ylab("Center of OpenPose Detection") + 
  ggtitle("Hand False Positives") + theme(legend.position = "none") +
  scale_y_reverse() +
  theme_few(base_size=12)

ggpubr::ggarrange(p1, p2)
```

@@

```{r roc_curves, fig.width=6.5, fig.height=3.2, fig.env="figure", fig.cap="ROC curves relating the rate of true vs. false positives for both face and hand detections at each level of detector confidence thresholds (lighter values = stricter cutoffs)."}
ggpubr::ggarrange(face_roc, hand_roc, nrow=1)
```

@@

ggpubr::ggarrange(p1, p2, nrow=1)
```

@@

ggpubr::ggarrange(p1, p2, nrow=1)
```

@@

@@

@@

```{r, fig.width=6.5, fig.height=3.0, fig.cap="Counts of activities by location in videos taking place in only a single location."}
# single location videos
meta_long_1loc %>% 
  mutate(Location = factor(Location, levels=names(loc_freq))) %>%
  group_by(Activity, Location) %>%
  summarise(n = n()) %>% 
  ggplot(aes(x=Activity, y=Location, fill=n)) +
  geom_tile() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
#ggsave("activity_by_location.pdf", width=6.5, height=3.5)
```

@@

ggplot(all_vid_bbs %>% filter(label %in% c('face','hand')), aes(x=avg_center_x , y = avg_center_y, size=num_detect, color=label)) +
  geom_point(alpha=.3) + 
  stat_density2d(aes(fill = label, color=label), geom="polygon", alpha=.6) +
  coord_fixed(ratio=.76) +
  # ggtitle('Face/hand density') +
  ylab('') +
  xlab('') +
  theme_few(base_size=14) +
  theme(legend.position="top", axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) +
  scale_y_reverse() +
  facet_wrap(~age_day_bin) +
  guides(size = FALSE) +
  ggthemes::scale_color_solarized(name = "")  +
  ggthemes::scale_fill_solarized(name = "") +
  theme(legend.position = c(0.8, 0.1))
```

>>>>

```{r handLocation, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=6.5, fig.height=3.5, fig.cap = "Density estimates for the child (left) and adult (right) hands that were detected in the 24K frame random gold set; each dot represents the center of a bounding box made by an adult participant. Brighter values indicate more detections."}
cowplot::plot_grid(child_hand_plot, adult_hand_plot, nrow=1)

ggsave('appendix_figs/handdensity.pdf', width = 7, height = 3)
```

@@

# density plot of OP's false positive face/hand locations
p1 <- ddd %>% filter(face_eval_ketan=="FP", label=="face") %>%
  mutate(x=left+0.5*width,
         y=1-(top+0.5*height)) %>% # could actually overplot all the face rectangles? prob very messy...
  ggplot(aes(x=x, y=y, fill=..level..)) + 
  stat_density_2d(geom = "polygon") + xlim(-.05,1.05) + ylim(-.05,1.0) + 
theme_few(base_size=12) +
  xlab("Center of OpenPose Detection") + 
  ylab("Center of OpenPose Detection") + 
  ggtitle("Face False Positives") + theme(legend.position = "none") +
  scale_y_reverse()

p2 <- ddd %>% filter(hand_eval_ketan=="FP", label=="hand") %>%
  mutate(x=left+0.5*width,
         y=1-(top+0.5*height)) %>% # top-left is 0,0
  ggplot(aes(x=x, y=y, fill=..level..)) + 
  stat_density_2d(geom = "polygon") + xlim(-.05,1.05) + ylim(-.05,1.0) + 
   theme_few(base_size=12) +
  xlab("Center of OpenPose Detection") + 
  ylab("Center of OpenPose Detection") + 
  ggtitle("Hand False Positives") + theme(legend.position = "none") +
  scale_y_reverse()

ggpubr::ggarrange(p1, p2)
ggsave('appendix_figs/op_falsepositives.pdf', width = 7, height = 3)

@@

```{r roc_curves, fig.width=6.5, fig.height=3.2, fig.env="figure", fig.cap="ROC curves relating the rate of true vs. false positives for both face and hand detections at each level of detector confidence thresholds (lighter values = stricter cutoffs)."}
ggpubr::ggarrange(face_roc, hand_roc, nrow=1)
ggsave('appendix_figs/rocs.pdf', width = 7, height = 3)
```

@@

ggpubr::ggarrange(p1, p2, nrow=1)
ggsave('appendix_figs/fscore_by_age_child.pdf', width = 7, height = 5)
```

@@

ggpubr::ggarrange(p1, p2, nrow=1)
ggsave('appendix_figs/fscore_by_location.pdf', width = 7, height = 3)
```

@@

ggsave('appendix_figs/humans_vs_op.pdf', width = 7, height = 5)

@@

ggsave('appendix_figs/activities_by_location.pdf', width = 7, height = 4)

@@

```{r, fig.width=6.5, fig.height=3.0, fig.cap="Counts of activities by location in videos taking place in only a single location."}
# single location videos
meta_long_1loc %>% 
  mutate(Location = factor(Location, levels=names(loc_freq))) %>%
  group_by(Activity, Location) %>%
  summarise(n = n()) %>% 
  ggplot(aes(x=Activity, y=Location, fill=n)) +
  geom_tile() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave('appendix_figs/activities_by_single_location.pdf', width = 7, height = 4)

@@

ggplot(all_vid_bbs %>% filter(label %in% c('face','hand')), aes(x=avg_center_x , y = avg_center_y, size=num_detect, color=label)) +
  geom_point(alpha=.3) + 
  stat_density2d(aes(fill = label, color=label), geom="polygon", alpha=.6) +
  coord_fixed(ratio=.76) +
  # ggtitle('Face/hand density') +
  ylab('') +
  xlab('') +
  theme_few(base_size=14) +
  theme(legend.position="top", axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) +
  scale_y_reverse() +
  facet_wrap(~age_day_bin) +
  guides(size = FALSE) +
  ggthemes::scale_color_solarized(name = "")  +
  ggthemes::scale_fill_solarized(name = "") +
  theme(legend.position = c(0.8, 0.1))

ggsave('appendix_figs/face_hand_density.pdf', width = 7, height = 7)
```

## A10 Face/hand detections by walking ability

Each parent in the SAYCam dataset completed the Ages and Stages Questionnaire (ASQ), which includes questions about their child's motor abilities, and we used the responses from this questionnaire to code the age (in months) at which each child became able to walk on their own.  We found considerable variability in the age at which each child achieved this milestone: S began walking after 10 months, A after 13 months, and Y after 16 months of age.

Below, we plot face and hand detections for each child as a function of when each child started walking independently.

```{r}
face_hand_by_age_milestones <- face_hand_by_age %>%
  filter(num_frames_total > 5000) %>%
  mutate(walking = ifelse((child_id=="S" & age_day_bin > 10) |
                       (child_id=="A" & age_day_bin > 13) |
                       (child_id=="Y" & age_day_bin > 16), 'Walking', 'Not Yet Walking')) %>%
  mutate(child_name = case_when(child_id =='A' ~ "Child A",
                      child_id == "S" ~ "Child S",
                      child_id == "Y" ~ "Child Y")) %>%
  select(prop_faces, prop_hands, num_frames_total, age_day_bin, child_name, walking) %>%
  pivot_longer(cols = c(prop_faces, prop_hands), names_to = "region", values_to = "prop")  %>%
    mutate(region = case_when(region =='prop_hands' ~ "Hands",
                      region == "prop_faces" ~ "Faces"))

```{r}
ggplot(face_hand_by_age_milestones,
       aes(x=age_day_bin, y=prop,
           size=log10(num_frames_total),
           col=walking)) +
  geom_point(alpha=.3) +
  ylab('Proportion Detections') +
  xlab('Age (Months)') +
  ylim(0,.5) +
  facet_grid(region ~ child_name) +
  theme_few(base_size=10) +
  viridis::scale_color_viridis(name = "", option="C", discrete=TRUE, begin=0, end=.4) +
  theme(legend.text=element_text(size=10)) +
  guides(size = FALSE) +
  theme(legend.position="right")

ggsave('face_hand_by_motor.pdf', width=7, height=5)
```

### 2022-01-07 paper/manuscript/manuscript_dec2021_apa_blinded.Rmd
save appendix pdf figs in line

---
title: A longitudinal analysis of the social information in infants' naturalistic
  visual experience using automated detections
author:
- name: Blinded
  affiliation: '1'
  corresponding: yes
  address: ''
  email: ' '
- name: ' '
  affiliation: '1'
shorttitle: Longitudinal analysis of social information
output: papaja::apa6_docx
authornote: "The data and code that support the findings of this study are available
  at https://osf.io/cdhw4/. \n"
abstract: |
  The faces and hands of caregivers and other social partners offer a rich source of social and causal information that is likely critical for infants' cognitive and linguistic development. Previous work using manual annotation strategies and cross-sectional data has found systematic changes in the proportion of faces and hands in the egocentric perspective of young infants. Here, we examine the prevalence of faces and hands in a longitudinal collection of more than 1700 headcam videos from three children ages 6 to 32 months. To analyze these naturalistic infant egocentric videos, we validated the use of a modern convolutional neural network (OpenPose) for the detection of faces and hands and then applied this model to the entire dataset. First, we found a higher proportion of hands in view than previously reported and a moderate decrease in the proportion of faces in children's view across age. Second, we found substantial variability in the proportion of faces and hands viewed by different children in different locations (e.g., living room vs. kitchen), suggesting that individual activity contexts may shape the social information that infants experience. Third, we found evidence that children may see closer, larger views of people, hands, and faces earlier in development. These analyses provide new insight into the changes in the social information in view across the first few years of life and call for further work that examines their generalizability across populations and their relationship to learning outcomes.
keywords: social cognition; face perception; infancy; head cameras; deep learning
wordcount: X
bibliography: library.bib
floatsintext: yes
figurelist: no
tablelist: no
footnotelist: no
linenumbers: no
mask: no
draft: no
documentclass: apa6
classoption: man
affiliation:
- id: '1'
  institution: Blinded

@@

Here, we examine the social information present in the infant visual perspective---the presence of faces and hands---by analyzing a longitudinal collection of more than 1700 headcam videos collected at home from three children along a span of 6 to 32 months of age—--the SAYCam dataset [@SAYcam]. We analyze the video data from this dense, longitudinal dataset as case studies of how the social information in the infant view changes over the course of development in individual children. These kinds of longitudinal case studies provide an important counterpoint to cross-sectional investigations for several reasons. By repeatedly recording data from the same child, they allow us to investigate developmental changes within each child, avoiding the ecological fallacy of assuming that between-person patterns reflect within-person mechanisms [@piantadosi1988ecological]. Second, they allow us to relate developmental changes to milestones achieved by an individual child (e.g., the onset of walking). Second, they eliminate several sources of variability (e.g., parent/child familiarity with camera, home environment) that may obscure fine-grained developmental patterns.

@@

We thus capitalize on this computational advance to annotate the entirety of this large, longitudinal video dataset. 
We use these detections to first report how the proportion of faces and hands changes with age in each of the three children in the dataset. 
We then investigate sources of variability in our more naturalistic dataset that may explain differences from prior work, including both the field-of-view of the head cameras as well as a diversity of locations in which videos were recorded.
Finally, making use of automated annotation of pose bounding boxes, we analyze the size, location, and variability of detected faces and poses across development in these three children.
Overall, these findings largely validate prior cross-sectional findings, and further suggest that automated annotated can be used to analyze naturalistic video datasets to build towards generalizable conclusions about the statistics of the social information in the infant view across a wide variety of populations.

@@

### Restricting detections to the center field-of-view
Given that the fish-eye lens may have distorted the edges of the visual field in the videos, we also examined whether we would see improved precision by restricting the included detections to the center field-of-view, and excluding detections that occured in the outer 20\% of each frame. This restricted field-of-view is also more comparable to video data from other head-mounted cameras. However, we found that excluding these detections did not improve the model's precision for either faces or hands; rather, it only decreased recall. Excluding peripehral detections led to an overall f-score of `r round(face_performance_center[3],2)`, for faces with a precision of `r round(face_performance_center[1],2)` and recall of `r round(face_performance_center[2],2)`. For hands, the F-score for center detections was `r round(hand_performance_center[3],2)`, with a precision of `r round(hand_performance_center[1],2)` and recall of `r round(hand_performance_center[2],2)`. Thus, detections that were made in the periphery of these videos were not more likely to be false alarms for either faces or hands. Appendix Figure A3 illustrates this finding, showing the density of false positives for both faces and hands that are concentrated in the center of the field of view.

@@

```{r}
first_panel <- ggplot(all_detections, 
       aes(x=age_day_bin, y=prop, 
           size=log10(num_frames_total),
           col=region)) +
  geom_point(alpha=.1) +
  geom_smooth(span=10, aes(weight = num_frames_total), show.legend = TRUE) + 
  ylab('Proportion Detections') + 
  xlab('Age (Months)') +
  ylim(0,.5) +
  facet_grid(cols = vars(child_name)) + 
  theme_few(base_size=10) +
  ggthemes::scale_color_solarized(name = "") + 
  theme(legend.text=element_text(size=10)) +
  guides(size = FALSE) +
  theme(legend.position="right")

second_panel <- ggplot(centered, 
       aes(x=age_day_bin, y=prop, 
           size=log10(num_frames_total),
           col=region)) +
  geom_point(alpha=.1) +
  geom_smooth(span=10, aes(weight = num_frames_total),linetype = 'dashed', show.legend = TRUE) + 
  ylab('Proportion Detections') + 
  xlab('Age (Months)') +
  ylim(0,.5) +
  facet_grid(cols = vars(child_name)) + 
  theme_few(base_size=10) +
  ggthemes::scale_color_solarized(name = "") + 
  theme(legend.text=element_text(size=10)) +
  guides(size = FALSE) +
  theme(legend.position="right")

third_panel <- ggplot(detailed, 
       aes(x=age_day_bin, y=prop, 
           size=log10(num_frames_total),
           col=region)) +
  geom_point(alpha=.1) +
  geom_smooth(span=10, aes(weight = num_frames_total), show.legend = TRUE) + 
  ylab('Proportion Detections') + 
  xlab('Age (Months)') +
  ylim(0,.5) +
  facet_grid(cols = vars(child_name)) + 
  theme_few(base_size=10) +
  scale_color_manual(values = c("#751098","#142c83"), name = "") +
  theme(legend.text=element_text(size=10)) +
  guides(size = FALSE) +
  theme(legend.position="right")

@@

```{r faceHandDetailed, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=6, fig.cap = "Proportion of frames with (A) All face and hand detections,  (B) Face/hand detections that fell within the center field-of-view (reducing the contribution of children's own hands) and (C) Face detections that were full faces (e.g., eyes, nose, and mouth all visible) and that co-occurred with hands, plotted as a function of age for each child (A, S, and Y). Data are binned by each week that the videos were filmed and scaled by the number of frames in that age range."}

cowplot::plot_grid(first_panel, second_panel, third_panel, labels='AUTO', nrow=3, align = "v", axis = "b")
```

>>>>

---
title: A longitudinal analysis of the social information in infants' naturalistic
  visual experience using automated detections
author:
- name: Blinded
  affiliation: '1'
  corresponding: yes
  address: ''
  email: ' '
- name: ' '
  affiliation: '1'
shorttitle: Longitudinal analysis of social information
output: papaja::apa6_docx
authornote: "The data and code that support the findings of this study are available
  at https://osf.io/cdhw4/. \n"
abstract: |
  The faces and hands of caregivers and other social partners offer a rich source of social and causal information that is likely critical for infants’ cognitive and linguistic development. Previous work using manual annotation strategies and cross-sectional data has found systematic changes in the proportion of faces and hands in the egocentric perspective of young infants. Here, we validated the use of a modern convolutional neural network (OpenPose) for the detection of faces and hands in naturalistic egocentric videos. We then applied this model to a longitudinal collection of more than 1700 headcam videos from three children ages 6 to 32 months. Using these detections, we confirm and extend prior results from cross-sectional studies. First, we found a moderate decrease in the proportion of faces in children’s view across age and a higher proportion of hands in view than previously reported. Second, we found variability in the proportion of faces and hands viewed by different children in different locations (e.g., living room vs. kitchen), suggesting that individual activity contexts may shape the social information that infants experience. Third, we found evidence that children may see closer, larger views of people, hands, and faces earlier in development. These longitudinal analyses provide an additional perspective on the changes in the social information in view across the first few years of life and suggest that pose detection models can successfully be applied to naturalistic egocentric video datasets to extract descriptives about infants’ changing social environment.
keywords: social cognition; face perception; infancy; head cameras; deep learning
wordcount: X
bibliography: library.bib
floatsintext: no
figurelist: no
tablelist: no
footnotelist: no
linenumbers: no
mask: no
draft: no
documentclass: apa6
classoption: man
affiliation:
- id: '1'
  institution: Blinded

@@

Here, we examine the social information present in the infant visual perspective -- the presence of faces and hands -- by analyzing a longitudinal collection of more than 1700 headcam videos collected at home from three children along a span of 6 to 32 months of age -- the SAYCam dataset [@SAYcam]. We analyze the video data from this dense, longitudinal dataset as case studies of how the social information in the infant view changes over the course of development in individual children. These kinds of longitudinal case studies provide an important counterpoint to cross-sectional investigations for several reasons. By repeatedly recording data from the same child, they allow us to investigate developmental changes within each child, avoiding the ecological fallacy of assuming that between-person patterns reflect within-person mechanisms [@piantadosi1988ecological]. Second, they allow us to relate developmental changes to milestones achieved by an individual child (e.g., the onset of walking). Third, they eliminate several sources of variability (e.g., parent/child familiarity with camera, home environment) that may obscure fine-grained developmental patterns.

@@

We thus capitalize on this computational advance to annotate the entirety of this large, longitudinal video dataset. 
We use these detections to first report how the proportion of faces and hands changes with age in each of the three children in the dataset. 
We then investigate sources of variability in our more naturalistic dataset that may explain differences from prior work, including both the field-of-view of the head cameras as well as a diversity of locations in which videos were recorded.
Finally, making use of automated annotation of pose bounding boxes, we analyze the size, location, and variability of detected faces and poses across development in these three children.
Overall, these findings largely validate prior cross-sectional findings, and further suggest that automated annotations can be used to analyze naturalistic video datasets to build towards generalizable conclusions about the statistics of the social information in the infant view across a wide variety of populations.

@@

### Restricting detections to the center field-of-view
Given that the fish-eye lens may have distorted the edges of the visual field in the videos, we also examined whether we would see improved precision by restricting the included detections to the center field-of-view, and excluding detections that occured in the outer 20\% of each frame. This restricted field-of-view is also more comparable to video data from other head-mounted cameras. However, we found that excluding these detections did not improve the model's precision for either faces or hands; rather, it only decreased recall. Excluding peripheral detections led to an overall f-score of `r round(face_performance_center[3],2)`, for faces with a precision of `r round(face_performance_center[1],2)` and recall of `r round(face_performance_center[2],2)`. For hands, the F-score for center detections was `r round(hand_performance_center[3],2)`, with a precision of `r round(hand_performance_center[1],2)` and recall of `r round(hand_performance_center[2],2)`. Thus, detections that were made in the periphery of these videos were not more likely to be false alarms for either faces or hands. Appendix Figure A3 illustrates this finding, showing the density of false positives for both faces and hands that are concentrated in the center of the field of view.

@@

```{r}
first_panel <- ggplot(all_detections, 
       aes(x=age_day_bin, y=prop, 
           size=log10(num_frames_total),
           col=region)) +
  geom_point(alpha=.1) +
  geom_smooth(span=10, aes(weight = num_frames_total), show.legend = TRUE) + 
  ylab('Proportion Detections') + 
  xlab('Age (Months)') +
  ylim(0,.5) +
  facet_grid(cols = vars(child_name)) + 
  theme_few(base_size=10) +
  ggthemes::scale_color_solarized(name = "") + 
  theme(legend.text=element_text(size=10)) +
  guides(size = FALSE) +
  theme(legend.position="bottom") +
  ggtitle('A.')

second_panel <- ggplot(centered, 
       aes(x=age_day_bin, y=prop, 
           size=log10(num_frames_total),
           col=region)) +
  geom_point(alpha=.1) +
  geom_smooth(span=10, aes(weight = num_frames_total),linetype = 'dashed', show.legend = TRUE) + 
  ylab('Proportion Detections') + 
  xlab('Age (Months)') +
  ylim(0,.5) +
  facet_grid(cols = vars(child_name)) + 
  theme_few(base_size=10) +
  ggthemes::scale_color_solarized(name = "") + 
  theme(legend.text=element_text(size=10)) +
  guides(size = FALSE) +
  theme(legend.position="bottom")  +
  ggtitle('B.')

third_panel <- ggplot(detailed, 
       aes(x=age_day_bin, y=prop, 
           size=log10(num_frames_total),
           col=region)) +
  geom_point(alpha=.1) +
  geom_smooth(span=10, aes(weight = num_frames_total), show.legend = TRUE) + 
  ylab('Proportion Detections') + 
  xlab('Age (Months)') +
  ylim(0,.5) +
  facet_grid(cols = vars(child_name)) + 
  theme_few(base_size=10) +
  scale_color_manual(values = c("#751098","#142c83"), name = "") +
  theme(legend.text=element_text(size=10)) +
  guides(size = FALSE) +
  theme(legend.position="bottom") +
  ggtitle('C.')

@@

```{r faceHandDetailed, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=7, fig.height=9, fig.cap = "Proportion of frames with (A) All face and hand detections,  (B) Face/hand detections that fell within the center field-of-view (reducing the contribution of children's own hands) and (C) Face detections that were full faces (e.g., eyes, nose, and mouth all visible) and that co-occurred with hands, plotted as a function of age for each child (A, S, and Y). Data are binned by each week that the videos were filmed and scaled by the number of frames in that age range."}

cowplot::plot_grid(first_panel, second_panel, third_panel, labels=NULL, nrow=3, label_fontface = "bold", align = "v", axis = "b")
ggsave('figs_pdfs/test.pdf', width = 7, height = 9)
```
